---
title: "Replicate 'data-ui' Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replicate 'data-ui' Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Install/Load dataui

`dataui` is currently in pre-alpha state with a potentially unstable API.  For now `dataui` will only live on Github, but cRAN is the ultimate goal.

```{r setup}
# remotes::install_github("timelyportfolio/dataui")
library(dataui)
```

## Sparkline Kitchen Sink Examples

Let's attempt to replicate the examples from `data-ui` [storybook](https://williaster.github.io/data-ui/?selectedKind=sparkline&selectedStory=Kitchen%20sink&full=0&addons=0&stories=1&panelRight=0).  For reference, the JavaScript code is on [Github](https://github.com/williaster/data-ui/blob/master/packages/demo/examples/03-sparkline/KitchenSinkExamples.jsx).

Unfortunately we cannot recreate the emoji label due to `htmlwidgets::JS()` scoping issues.  This likely would be solved by [pull](https://github.com/rstudio/htmltools/pull/72) but no one seems to care.

```{r kitchensink1}
# unable to add smiley face due to scoping issues with htmlwidgets::JS :(
dui_sparkline_with_tooltip(
  # don't have to use reactR::babel_transform but makes it easy to copy/paste
  renderTooltip = htmlwidgets::JS(htmltools::HTML(reactR::babel_transform(
"
(
  { datum }, // eslint-disable-line react/prop-types
) => {
return (
  <div>
    {datum.x && <div>{datum.x}</div>}
    <div>{datum.y ? datum.y.toFixed(2) : '--'}</div>
  </div>
)}
"
  ))),
  data = lapply(1:35, function(i) {
    list(
      x = paste0("Day ", i),
      y = runif(1, 0, 2)
    )
  }),
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparkbarseries(fill = scales::brewer_pal(palette="Purples")(3)[2])
  )
)
```

```{r kitchensink2}
dui_sparkline_with_tooltip(
  data = lapply(1:25, function(i) {
    list(
      x = paste0("Day ", i),
      y = runif(1, 0, 2)
    )
  }),
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparkpatternlines(
      id = "area_pattern",
      height = 4,
      width = 4,
      stroke = "#91a7ff",
      strokeWidth = 1,
      orientation = "diagonal"
    ),
    dui_sparklineseries(
      showArea = TRUE,
      stroke = "#91a7ff",
      fill = "url(#area_pattern)"
    ),
    dui_sparkpointseries(
      points = list("all"),
      stroke = "rgb(92, 124, 250)",
      fill = "#fff",
      size = 3
    ),
    dui_sparkpointseries(
      points = list("last"),
      fill = "rgb(92, 124, 250)",
      renderLabel = htmlwidgets::JS("d => d.toFixed(2)"),
      labelPosition = "right"
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          key = "ref-line",
          strokeWidth = 1,
          strokeDasharray = "4 4"
        ),
        dui_sparkpointseries(
          key = "ref-point",
          fill = "rgb(92, 124, 250)"
        )
      )
    )
  )
)
```

```{r kitchensink3}
dui_sparkline_with_tooltip(
  data = lapply(1:25, function(i) {
    list(
      x = paste0("Day ", i),
      y = runif(1, 0, 2)
    )
  }),
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparklineseries(
      stroke = "#f06595"
    ),
    dui_sparkpointseries(
      points = list("all"),
      stroke = "#f06595",
      fill = "#f06595",
      size = 3
    ),
    dui_sparkpointseries(
      points = list("min","max"),
      fill = "#fcc419",
      stroke = "#fff",
      renderLabel = htmlwidgets::JS("d => d.toFixed(2)"),
      labelPosition = "right",
      size = 5
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          key = "ref-line",
          strokeWidth = 1,
          strokeDasharray = "4 4"
        ),
        dui_sparkpointseries(
          key = "ref-point",
          fill = "#fff",
          stroke = "#f06595"
        )
      )
    )
  )
)
```

```{r kitchensink4}
dui_sparkline_with_tooltip(
  data = lapply(1:25, function(i) {
    list(
      x = paste0("Day ", i),
      y = runif(1, 0, 2)
    )
  }),
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparklineseries(
      stroke = "#1098ad"
    ),
    dui_sparkhorizontalrefline(
      stroke = "#1098ad",
      strokeWidth = 1,
      strokeDasharray = "4 4"
    ),
    dui_sparkpointseries(
      points = list("first","last"),
      fill = "#1098ad",
      stroke = "#fff",
      renderLabel = htmlwidgets::JS("d => d.toFixed(2)"),
      labelPosition = htmlwidgets::JS("{(d, i) => (i === 0 ? 'left' : 'right')}"),
      size = 5
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          key = "ref-line",
          strokeWidth = 1,
          strokeDasharray = "4 4"
        ),
        dui_sparkpointseries(
          key = "ref-point",
          stroke = "#fff",
          fill = "#fcc419"
        )
      )
    )
  )
)
```

```{r kitchensink5}
dui_sparkline_with_tooltip(
  data = lapply(1:25, function(i) {
    list(
      x = paste0("Day ", i),
      y = runif(1, 0, 2)
    )
  }),
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparkpatternlines(
      id = "band_pattern",
      height = 6,
      width = 6,
      stroke = "#be4bdb",
      strokeWidth = 1,
      orientation = list("diagonal")
    ),
    dui_sparkbandline(fill = "url(#band_pattern)"),
    dui_sparklineseries(
      stroke = "#be4bdb"
    ),
    dui_sparkhorizontalrefline(
      stroke = "#862e9c",
      strokeWidth = 1,
      strokeDasharray = "4 4",
      reference = "median"
    ),
    dui_sparkverticalrefline(
      stroke = "#e599f7",
      strokeDasharray = "4 4",
      reference = "min"
    ),
    dui_sparkverticalrefline(
      stroke = "#e599f7",
      strokeDasharray = "4 4",
      reference = "max"
    ),
    dui_sparkpointseries(
      points = list("min","max"),
      fill = "#da77f2",
      stroke = "#fff",
      renderLabel = htmlwidgets::JS("d => d.toFixed(2)"),
      labelPosition = htmlwidgets::JS("{(d, i) => (i === 0 ? 'left' : 'right')}"),
      size = 5
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          key = "ref-line",
          strokeWidth = 1,
          strokeDasharray = "4 4"
        ),
        dui_sparkpointseries(
          key = "ref-point",
          stroke = "#fff",
          fill = "#fcc419"
        )
      )
    )
  )
)
```

```{r kitchensink6}
dat <- lapply(1:35, function(i) {
  list(
    x = paste0("Day ", i),
    y = runif(1, 0, 2)
  )
})

dui_sparkline_with_tooltip(
  data = dat,
  valueAccessor = htmlwidgets::JS("d => d.y"),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 64 ),
  ariaLabel = "sparkline example",
  components = list(
    dui_sparkpatternlines(
      id = "band_pattern_hash",
      height = 7,
      width = 7,
      stroke = "#ced4da",
      strokeWidth = 1,
      orientation = list("vertical", "horizontal")
    ),
    dui_sparkbandline(
      fill = "url(#band_pattern_hash)",
      stroke = "#ced4da",
      band = list(
        from = min(unlist(lapply(dat, function(d) d$y))),
        to = max(unlist(lapply(dat, function(d) d$y)))
      ) # data-ui code not working so replicate in R with min and max
    ),
    dui_sparklineseries(
      stroke = "#495057"
    ),
    # add these for fun from prior example
    dui_sparkhorizontalrefline(
      stroke = "#343a40",
      strokeWidth = 1,
      strokeDasharray = "4 4",
      reference = "median"
    ),
    dui_sparkverticalrefline(
      stroke = "#e599f7",
      strokeDasharray = "4 4",
      reference = "min"
    ),
    dui_sparkverticalrefline(
      stroke = "#e599f7",
      strokeDasharray = "4 4",
      reference = "max"
    ),
    dui_sparkpointseries(
      points = list("min","max"),
      fill = "#ced4da",
      stroke = "#fff",
      renderLabel = htmlwidgets::JS("d => d.toFixed(2)"),
      labelPosition = htmlwidgets::JS("{(d, i) => (i === 0 ? 'left' : 'right')}"),
      size = 5
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          key = "ref-line",
          strokeWidth = 1,
          strokeDasharray = "4 4"
        ),
        dui_sparkpointseries(
          key = "ref-point",
          stroke = "#fff",
          fill = "#343a40"
        )
      )
    )
  )
)
```

## Sparkline Bar Examples

We were almost able to fully replicate the kitchen sink examples.  Now, let's look at the  examples from `data-ui` [sparkline bar storybook](https://williaster.github.io/data-ui/?selectedKind=sparkline&selectedStory=BarSeries&full=0&addons=0&stories=1&panelRight=0).  For reference, the JavaScript code is on [Github](https://github.com/williaster/data-ui/blob/master/packages/demo/examples/03-sparkline/BarSeriesExamples.jsx).


```{r bar1}
dui_sparkline(
  data = runif(50, 0, 2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 8 ),
  ariaLabel = "sparkline bar example",
  components = list(
    dui_sparkbarseries()
  )
)
```

```{r bar2}
dui_sparkline_with_tooltip(
  data = runif(35, 0, 5) + 1:35 ,
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 8 ),
  ariaLabel = "sparkline bar example",
  components = list(
    dui_sparkbarseries(
      fillOpacity = 0.7,
      fill = "#eebefa"
    ),
    dui_tooltip(
      components = list(
        dui_sparkhorizontalrefline(
          stroke = "#9c36b5",
          strokeWidth = 1,
          strokeDasharray = "3,3",
          # not able to recreate label due to htmlwidgets::JS scope
          labelPosition = "right",
          labelOffset = "12",
          renderLabel = htmlwidgets::JS("d => d.toFixed(1)")
        )
      )
    )
  )
)
```

```{r bar3}
dui_sparkline_with_tooltip(
  data = runif(30, 0, 2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 8 ),
  ariaLabel = "sparkline bar example",
  components = list(
    dui_sparklineargradient(
      id = "bar_gradient",
      to = "#fff3bf",
      from = "#f08c00"
    ),
    dui_sparkbarseries(
      fillOpacity = htmlwidgets::JS("(d, i) => (i === 24 ? 1 : 0.5)"),
      fill = "url(#bar_gradient)"
    ),
    dui_sparkverticalrefline(
      reference = 24,
      stroke = "#f08c00",
      strokeWidth = 1,
      strokeDasharray = "3,3",
      renderLabel = htmlwidgets::JS("d => d.toFixed(1)")
    ),
    dui_tooltip(
      components = list(
        dui_sparkhorizontalrefline(
          stroke = "#f08c00",
          strokeWidth = 1,
          strokeDasharray = "3,3",
          renderLabel = htmlwidgets::JS("d => d.toFixed(1)")
        )
      )
    )
  )
)
```

```{r bar4}
dui_sparkline_with_tooltip(
  data = runif(30, 0, 2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 8 ),
  ariaLabel = "sparkline bar example",
  components = list(
    dui_sparkbarseries(
      fillOpacity = 0.9,
      fill = "#63e6be"
    ),
    dui_sparklineseries(
      stroke = "#087f5b"
    )
  )
)
```

```{r bar5}
dui_sparkline_with_tooltip(
  data = runif(20, 0, 2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 64, bottom= 24, left= 8 ),
  ariaLabel = "sparkline bar example",
  components = list(
    dui_sparkpatternlines(
      id = "bar_pattern",
      height = 4,
      width = 4,
      stroke = "#868e96",
      strokeWidth = 1,
      orientation = list('diagonal')
    ),
    dui_sparkpatternlines(
      id = "bar_pattern_2",
      height = 4,
      width = 4,
      stroke = "#d6336c",
      strokeWidth = 1,
      orientation = list('horizontal', 'vertical')
    ),
    dui_sparkbarseries(
      fill = htmlwidgets::JS(htmltools::HTML(
        "(d, i) => (i > 9 ? 'url(#bar_pattern_2)' : 'url(#bar_pattern)')"
      )),
      strokeWidth = 5
    ),
    dui_sparkverticalrefline(
      reference = 9.5,
      stroke = "#343a40",
      strokeWidth = 1,
      strokeDasharray = "2,2",
      strokeLinecap = "square",
      # let's not use reactR::babel_transform this time
      renderLabel = htmlwidgets::JS(htmltools::HTML("
() => React.createElement('tspan', {fill: '#d6336c'}, '!!')
      "))
    )
  )
)
```


## Sparkline Line Examples

Since we are making good progress, we can now explore the examples from `data-ui` [sparkline line storybook](https://williaster.github.io/data-ui/?selectedKind=sparkline&selectedStory=LineSeries&full=0&addons=0&stories=1&panelRight=0).  For reference, the JavaScript code is on [Github](https://github.com/williaster/data-ui/blob/master/packages/demo/examples/03-sparkline/LineSeriesExamples.jsx).

```{r line1}
dui_sparkline(
  data = runif(50,0,2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 90, bottom= 24, left= 8 ),
  ariaLabel = "sparkline line example",
  components = list(
    dui_sparklineseries(),
    dui_sparkpointseries(
      points = list('last'),
      size = 4,
      stroke = "white",
      renderLabel = htmlwidgets::JS("d => d.toFixed(1)"),
      labelPosition = "right"
    )
  )
)
```

```{r line2}
dui_sparkline_with_tooltip(
  data = runif(40,0,2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 90, bottom= 24, left= 8 ),
  ariaLabel = "sparkline line example",
  components = list(
    dui_sparklineargradient(
      id = "area_gradient",
      to = "#ffdeeb",
      from = "#e64980"
    ),
    dui_sparklineseries(
      showArea = TRUE,
      fill = "url(#area_gradient)",
      stroke = "#f783ac"
    ),
    dui_tooltip(
      components = list(
        dui_sparkverticalrefline(
          strokeWidth = 1,
          stroke = "#e64980",
          strokeDasharray="4 4"          
        ),
        dui_sparkpointseries(
          fill = "#e64980"
        )
      )
    )
  )
)
```

```{r line3}
dui_sparkline_with_tooltip(
  data = runif(30,0,2),
  width = 500,
  height = 100,
  margin = list( top= 24, right= 90, bottom= 24, left= 8 ),
  ariaLabel = "sparkline line example",
  components = list(
    dui_sparkpatternlines(
      id = "innerquartile_pattern",
      height = 4,
      width = 4,
      stroke = "#329af0",
      strokeWidth = 1,
      orientation= list("diagonal")
    ),
    dui_sparkbandline(
      band = "innerquartiles",
      fill = "url(#innerquartile_pattern)"
    ),
    dui_sparklineseries(
      curve = "linear",
      stroke = "#4dadf7"
    ),
    dui_sparkhorizontalrefline(
      stroke = "#329af0",
      strokeDasharray = "3,3",
      strokeLinecap = "square",
      strokeWidth = 1,
      renderLabel = htmlwidgets::JS("d => `Median ${d.toFixed(1)}`")
    ),
    dui_sparkpointseries(
      points = list('min', 'max'),
      fill = "#228ae6",
      size = 4,
      stroke = "white",
      renderLabel = htmlwidgets::JS("d => d.toFixed(1)")
    )
  )
)
```

```{r line4}
curvesData <-  c(-10, 10, -10, 5, -5, 0)
curves <-  c('cardinal', 'linear', 'monotoneX', 'basis')

dui_sparkline(
  data = curvesData,
  width = 500,
  height = 100,
  margin = list( top= 24, right= 90, bottom= 24, left= 8 ),
  ariaLabel = "sparkline line example",
  components = c(
    mapply(
      function(d,i) {
        dui_sparkverticalrefline(
          key = i,
          reference = i,
          stroke = "#dee2e6",
          strokeWidth = 1,
          strokeLineCap = "square",
          strokeDasharray = "2,3",
          labelPosition = "right"
        )
      },
      curvesData,
      seq_along(curvesData) - 1,
      SIMPLIFY = FALSE,
      USE.NAMES = FALSE
    ),
    mapply(
      function(curve,i) {
        dui_sparklineseries(
          key = curve,
          curve = curve,
          stroke = blues9[i+2],
          strokeWidth = 2
        )
      },
      curves,
      seq_along(curves) - 1,
      SIMPLIFY = FALSE,
      USE.NAMES = FALSE
    ),
    mapply(
      function(curve,i) {
        dui_sparklabel(
          key = curve,
          textAnchor = "start",
          x = 500 - 90,
          dy = 8 + i * 12,
          label = curve,
          fill = blues9[i+2],
          strokeWidth = 1
        )
      },
      curves,
      seq_along(curves) - 1,
      SIMPLIFY = FALSE,
      USE.NAMES = FALSE
    )
  )
)
```
